Trauma treatment delay

### 2023-11-08 manuscriptlongversion.Rmd
code

```{r}
  factors.data %>%
    tbl_summary(  # Create summary statistics for different groups
      by = OFI_delay,
      label = list(
        Age = "Age (years)",
        Gender = "Gender",
        Total_GCS = "GCS"
      ),
      statistic = list(all_continuous() ~ "{mean} ± {sd}"),
      digits = list(all_continuous() ~ c(2, 2))
    ) %>%
    add_p()  # Add p-values

```

>>>>

![](table1.png)

Out of the 11,864, 5,282 patients were not screened for ofi, which
excludes them from the study. Out of the remaining 6,582 patients, 4,775
patients had missing data in one or more of the analysed parameters,
which led to exclusion from the study.

In the remaining group of 1,835 patients who have been screened for ofi
and have data registered in all parameters examined for the study, 28
patients were found to have received delayed treatment during their stay
at the hospital.

![](Rplot.png)

In the following table, the results from the regression analysis is
displayed.

The odds ratio is calculated, as well as a 95% confidence interval. The
p value is also shown.

### 2023-11-08 manuscriptlongversion.tex
code

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Trauma treatment delay},
  pdfauthor={Anton Wasielewski},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

@@

\hypertarget{ethical-considerations}{%
\subsection*{Ethical considerations}\label{ethical-considerations}}
\addcontentsline{toc}{subsection}{Ethical considerations}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Trauma treatment delay},
  pdfauthor={Anton Wasielewski},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Rstudio was used to perform the logistical regression. The dependent
variable is OFI\_delay, which is the presence of the ofi delayed
treatment.

\hypertarget{ethical-considerations}{%
\subsection{Ethical considerations}\label{ethical-considerations}}

Ethical permit is required and exists. The ethical review number is
2021-02541 and 2021-03531. An ethical problem that might be brought up
is the anonymity of the patients. In the registry, patients social
security number is present and could theoretically be used to identify
the patient in question. This will not be a problem though, as in the
analysis, the social security number is removed from the data before the
logistical regression is ran. Furthermore, the results will be presented
with patients grouped together, not individually. This will make it
impossible for anyone to identify patients. During the analysis, a VPN
was used to access the database to ensure that no data was leaked or
used for other purposes.

\hypertarget{results}{%
\subsection{Results}\label{results}}

Out of the 11864 patients treated at KS between 2012 and 2022, 1835 were
included in the study. The patient characteristics were as follows in
Table 1.

@@

Out of the 11,864, 5,282 patients were not screened for ofi, which
excludes them from the study. Out of the remaining 6,582 patients, 4,775
patients had missing data in one or more of the analysed parameters,
which led to exclusion from the study.

In the remaining group of 1,835 patients who have been screened for ofi
and have data registered in all parameters examined for the study, 28
patients were found to have received delayed treatment during their stay
at the hospital.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(Gmisc, }\AttributeTok{quietly =} \ConstantTok{TRUE}\NormalTok{)}
\FunctionTok{library}\NormalTok{(glue)}
\FunctionTok{library}\NormalTok{(htmlTable)}
\FunctionTok{library}\NormalTok{(grid)}
\FunctionTok{library}\NormalTok{(magrittr)}
\FunctionTok{library}\NormalTok{(gtsummary)}

\NormalTok{treated }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Treated at KS between 2012 and 2022"}\NormalTok{,}
                           \StringTok{"n = \{pop\}"}\NormalTok{,}
                           \AttributeTok{pop =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{11864}\NormalTok{),}
                           \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}
\NormalTok{screened }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Screened for ofi"}\NormalTok{,}
                         \StringTok{"n = \{pop\}"}\NormalTok{,}
                         \AttributeTok{pop =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{6582}\NormalTok{),}
                         \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}
\NormalTok{included }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Included in study"}\NormalTok{,}
                         \StringTok{"n = \{incl\}"}\NormalTok{,}
                         \AttributeTok{incl =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{1835}\NormalTok{),}
                         \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}
\NormalTok{delay }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Delay in treatment"}\NormalTok{,}
                      \StringTok{"n = \{recr\}"}\NormalTok{,}
                      \AttributeTok{recr =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{28}\NormalTok{),}
                      \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}

\NormalTok{no\_delay }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"No delay in treatment"}\NormalTok{,}
                      \StringTok{"n = \{recr\}"}\NormalTok{,}
                      \AttributeTok{recr =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{1807}\NormalTok{),}
                      \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}

\NormalTok{not\_screened }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Not screened for ofi"}\NormalTok{,}
                             \StringTok{"n = \{recr\}"}\NormalTok{,}
                             \AttributeTok{recr =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{11864{-}6582}\NormalTok{),}
                             \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}

\NormalTok{missing\_data }\OtherTok{\textless{}{-}} \FunctionTok{boxGrob}\NormalTok{(}\FunctionTok{glue}\NormalTok{(}\StringTok{"Missing data"}\NormalTok{,}
                             \StringTok{"n = \{recr\}"}\NormalTok{,}
                             \AttributeTok{recr =} \FunctionTok{txtInt}\NormalTok{(}\DecValTok{6582{-}1807}\NormalTok{),}
                             \AttributeTok{.sep =} \StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{))}
\FunctionTok{grid.newpage}\NormalTok{()}
\NormalTok{vert }\OtherTok{\textless{}{-}} \FunctionTok{spreadVertical}\NormalTok{(}\AttributeTok{treated =}\NormalTok{ treated,}
                       \AttributeTok{screened =}\NormalTok{ screened,}
                       \AttributeTok{included =}\NormalTok{ included,}
                       \AttributeTok{grps =}\NormalTok{ delay)}
\NormalTok{grps }\OtherTok{\textless{}{-}} \FunctionTok{alignVertical}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ vert}\SpecialCharTok{$}\NormalTok{grps,}
\NormalTok{                      delay, no\_delay) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{spreadHorizontal}\NormalTok{()}
\NormalTok{vert}\SpecialCharTok{$}\NormalTok{grps }\OtherTok{\textless{}{-}} \ConstantTok{NULL}

\NormalTok{not\_screened }\OtherTok{\textless{}{-}} \FunctionTok{moveBox}\NormalTok{(not\_screened,}
                    \AttributeTok{x =}\NormalTok{ .}\DecValTok{85}\NormalTok{,}
                    \AttributeTok{y =}\NormalTok{ .}\DecValTok{77}\NormalTok{,)}

\NormalTok{missing\_data }\OtherTok{\textless{}{-}} \FunctionTok{moveBox}\NormalTok{(missing\_data,}
                    \AttributeTok{x =}\NormalTok{ .}\DecValTok{85}\NormalTok{,}
                    \AttributeTok{y =}\NormalTok{ .}\DecValTok{5}\NormalTok{,)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{(}\FunctionTok{length}\NormalTok{(vert) }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)) \{}
  \FunctionTok{connectGrob}\NormalTok{(vert[[i]], vert[[i }\SpecialCharTok{+} \DecValTok{1}\NormalTok{]], }\AttributeTok{type =} \StringTok{"vert"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{    print}
\NormalTok{\}}
\FunctionTok{connectGrob}\NormalTok{(vert}\SpecialCharTok{$}\NormalTok{included, grps[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{type =} \StringTok{"N"}\NormalTok{)}
\FunctionTok{connectGrob}\NormalTok{(vert}\SpecialCharTok{$}\NormalTok{included, grps[[}\DecValTok{2}\NormalTok{]], }\AttributeTok{type =} \StringTok{"N"}\NormalTok{)}
\FunctionTok{connectGrob}\NormalTok{(vert}\SpecialCharTok{$}\NormalTok{treated, not\_screened, }\AttributeTok{type =} \StringTok{"L"}\NormalTok{)}
\FunctionTok{connectGrob}\NormalTok{(vert}\SpecialCharTok{$}\NormalTok{screened, missing\_data, }\AttributeTok{type =} \StringTok{"L"}\NormalTok{)}

\CommentTok{\# Print boxes}
\NormalTok{vert}
\NormalTok{grps}
\NormalTok{not\_screened}
\NormalTok{missing\_data}
\end{Highlighting}
\end{Shaded}

\includegraphics{manuscriptlongversion_files/figure-latex/unnamed-chunk-1-1.pdf}

In the following table, the results from the regression analysis is
displayed.

The odds ratio is calculated, as well as a 95\% confidence interval. The
p value is also shown.

\includegraphics{results.html}
